---
output:
  html_document
params:
  temp_dir: ""
  set_title: ""
title: "`r params$set_title`"
# toc: true
# toc_float: true
# toc_collapsed: true
# toc_depth: 3
---


<!-- todo:  -->
<!-- - PIT histograms -->
<!-- toc -->

```{r setup, include=FALSE}
# packages
library(scoringutils)
library(data.table)
library(dplyr)
library(tibble)
library(here)
library(purrr)
library(readr)
library(magrittr)
library(knitr)
library(ggbump)

# knitr options
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      warning = FALSE,
                      out.width = "100%")

temp_dir <- params$temp_dir

# locations <- params$locations
# locations <- c("Germany", "Poland")
# forecast_dates <- params$forecast_dates
# horizons <- params$horizons

# target_types = c("case", "death")


```

```{r include = FALSE}

# define parameters
parameters <- readRDS(file = paste0(temp_dir, "/scoring_parameters.RDS"))
locations <- parameters$locations
forecast_dates <- parameters$forecast_dates
horizons <- parameters$horizons
target_types = parameters$target_types

truth_data <- readRDS(file = paste0(temp_dir, "/truth_data.RDS")) %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date))
prediction_data <- readRDS(file = paste0(temp_dir, "/prediction_data.RDS")) %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date),
                forecast_date = as.character(forecast_date))

data <- dplyr::full_join(truth_data, prediction_data)

  # if (locations[1] == "all") {
  #   locations <- prediction_data$location_name %>%
  #     unique()
  # }
  #
  # if (forecast_dates[1] == "all") {
  #   forecast_dates <- prediction_data %>%
  #     dplyr::pull(forecast_date) %>%
  #     unique() %>%
  #     as.Date() %>%
  #     as.character() # for some reason numbers appear if I don't do that
# }
#
# if (horizons[1] == "all") {
#   horizons <- prediction_data %>%
#     dplyr::pull(horizon) %>%
#     unique()
# }

```


``` {r helper-functions}
# filter data
filter_helper <- function(dataframe, forecast_date = NULL, location = NULL,
                          horizon = NULL, target_type = NULL) {
  if (!is.null(forecast_date)) {
    forecast_date <- forecast_date # why do I need this?
    dataframe <- dplyr::filter(dataframe,
                               forecast_date == {{forecast_date}})
  }
  if (!is.null(location)) {
    location <- location # and this?
    dataframe <- dplyr::filter(dataframe,
                               location_name == {{location}})
  }
  if (!is.null(horizon)) {
    horizon <- horizon
    dataframe <- dplyr::filter(dataframe,
                               horizon == {{horizon}})
  }
  if (!is.null(target_type)) {
    target_type <- target_type
    dataframe <- dplyr::filter(dataframe,
                               target_type == {{target_type}})
  }
  return(dataframe)
}

check_print <- function(plot) {
  if (!is.null(plot)) {
    print(plot)
    cat("\n\n")
  }
}

```

```{r introduction-text, results = 'asis'}
if (!is.null(parameters$intro_text)) {
  cat(paste0(parameters$intro_text))
}

# This is a preliminary evaluation of forecasts made for the [German Forecast Hub](https://github.com/KITmetricslab/covid19-forecast-hub-de). These evaluations are preliminary - this means I cannot currently rule out any mistakes and the plots and analyses are subject to change. The evaluations are not authorised by the German Forecast Hub team.
# 
# Feel free to reproduce this analysis. To that end, you can clone this repository and the Forecast Hub repository, specify the directory of the submissions files in the R-Markdown script and run the script.
# 
# If you have questions or want to give feedback, please create an issue in this repository.


```

# Available Forecasts {.tabset}

``` {r availability, results = 'asis', warning = FALSE, fig.width = 10, eval = TRUE}
plot <- scoringutils::show_avail_forecasts(prediction_data, 
                                           facet_formula = location ~ target_type)

print(plot)

```


# Forecast Visualisation {.tabset}

``` {r showdata, results = 'asis', warning = FALSE, fig.width = 10, eval = TRUE}
plot_predictions_helper <- function(data,
                                    forecast_date = NULL,
                                    location = NULL,
                                    target_type = NULL,
                                    n_weeks = 12) {

  df <- filter_helper(data, forecast_date, location, target_type = target_type)
  location <- location
  target_type <- target_type
  additional_data <- data %>%
    dplyr::filter(location_name == {{location}},
                  target_type == {{target_type}},
                  target_end_date >= (Sys.Date() - n_weeks * 7)) %>%
    dplyr::select(-model)

  if (nrow(df) == 0) {
    return (NULL)
  }

  # maybe switch to using only one data source and filtering NA values
  plot <- scoringutils::plot_predictions(df,
                                         x = "target_end_date",
                                         facet_formula = ~ model,
                                         add_truth_data = additional_data,
                                         scales = "free") +
    ggplot2::ggtitle(paste("Predictions for incident", target_type,  "in", location))

  return(plot)
}

for (location in locations) {
  cat("\n\n##", location, "{.tabset}\n\n")
  for (forecast_date in forecast_dates) {
    cat(paste0("### ", forecast_date), "\n\n")
    for (target_type in target_types) {
      # cat(paste0("#### ", target_type), "\n\n")
      p1 <- plot_predictions_helper(data = data,
                                    forecast_date = forecast_date,
                                    location = location,
                                    target_type = target_type)
      check_print(p1)
    }
  }
}
```


# Ranking by pairwise comparison

``` {r pairwise, results = 'asis', warning = FALSE, fig.width = 10, eval = TRUE}

unsummarised_scores <- scoringutils::eval_forecasts(data)

# print(unsummarised_scores)

pairwise <- scoringutils::pairwise_comparison(unsummarised_scores)

# print(pairwise)

plot1 <- scoringutils::plot_pairwise_comparison(pairwise, type = "ratio")

print(plot1)

plot2 <- scoringutils::plot_pairwise_comparison(pairwise, type = "pval")
print(plot2)

```



# Score Ranking {.tabset}


Only those forecasters are named who have submitted a forecast in the most recent evaluated week.


``` {r results = 'asis', eval = TRUE}
plot_bump_helper <- function(data, target_type = NULL, location = NULL,
                             forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)

  if (nrow(df) == 0) {
    return (NULL)
  }

  scores <- scoringutils::eval_forecasts(df,
                                         summarise_by = c("model", "forecast_date") )
  if (nrow(scores) == 0) {
    return(NULL)
  }
  scores <- scores %>%
    dplyr::mutate(forecast_date = as.Date(forecast_date)) %>%
    dplyr::select(-coverage, -quantile_coverage) %>%
    dplyr::group_by(forecast_date) %>%
    dplyr::mutate(rank = rank(interval_score, ties.method = "average",
                              na.last = NA)) %>%
    dplyr::ungroup()

  plot <- ggplot2::ggplot(scores,
                          ggplot2::aes(x = forecast_date,
                                       y = rank, color = model)) +
    ggplot2::geom_point(size = 3, alpha = 1) + 
    ggbump::geom_bump(smooth = 15, size = 3, alpha = 0.1, na.rm = TRUE) +
    ggbump::geom_bump(smooth = 15, size = 0.5, alpha = 1, na.rm = TRUE) +
    ggplot2::geom_text(data = scores %>%
                         dplyr::filter(forecast_date == max(forecast_date)),
                       ggplot2::aes(label = model),
                       color = "grey10",
                       nudge_x = .61,
                       hjust = 0,
                       size = 2.6) +
    ggplot2::theme_minimal() +
    ggplot2::theme(legend.position = "none") +
    ggplot2::scale_y_reverse() +
    ggplot2::scale_x_date(expand = ggplot2::expansion(add = c(0, 12))) +
    ggplot2::labs(title = paste("Incident", target_type, "forecasts"))
  

  

  # plot <- scoringutils::score_table(scores) +
  #   ggplot2::ggtitle(paste("Incident", target_type, "forecasts"))
  return(plot)
}

plot_function <- plot_bump_helper

# replace by template ##########################################################
cat("\n\n## Overall \n")
pp <- plot_function(data)
check_print(pp)

if (length(target_types) > 1) {
  cat("\n\n## Overall separate \n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}

# by location
for (location in locations) {
  cat("\n\n##", location, "\n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type, location = location)
    check_print(p1)
  }
}
# replace by template ##########################################################



```





# Score Overview {.tabset .tabset-pills}

Overview of all the models and metrics

``` {r}
plot_score_table_helper <- function(data, target_type = NULL, location = NULL,
                                    forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)

  if (nrow(df) == 0) {
    return (NULL)
  } else {
    scores <- scoringutils::eval_forecasts(df,
                                           summarise_by = "model")

    if (nrow(scores) == 0) {
      return(NULL)
    }
    else {
      scores <- scores %>%
        dplyr::select(-coverage, -quantile_coverage)
    }

    plot <- scoringutils::score_table(scores) +
      ggplot2::ggtitle(paste("Incident", target_type, "forecasts"))
    return(plot)
  }
}
```

``` {r scoretable, fig.width = 12, results = 'asis', eval = TRUE}
# overall summary

plot_function <- plot_score_table_helper

# replace by template ##########################################################
cat("\n\n## Overall \n\n")
pp <- plot_function(data)
check_print(pp)

if (length(target_types) > 1) {
  cat("\n\n## Overall separate \n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}

# by location
for (location in locations) {
  cat("\n\n##", location, "{.tabset}\n\n")

  cat("\n\n### Overall \n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type, location = location)
    check_print(p1)
  }


  for (forecast_date in forecast_dates) {
    cat(paste0("\n\n### ", forecast_date), "\n\n")
    for (target_type in target_types) {
      p1 <- plot_function(data, target_type = target_type,
                          location = location,
                          forecast_date = forecast_date)
      check_print(p1)
    }
  }
}
# replace by template ##########################################################

```

# Overiew of Scores by Horizon {.tabset .tabset-pill}

Overview of all models and metrics stratified by forecast horizons

``` {r scoretablehorizon, fig.width = 12, results = 'asis', eval = TRUE}
# replace by template ##########################################################

plot_function <- plot_score_table_helper

for (horizon in horizons) {
  cat("\n\n## horizon = ", horizon, "{.tabset}\n\n")

  cat("\n\n### Overall\n\n")
  pp <- plot_function(data, horizon = horizon)
  check_print(pp)

  if (length(target_types) > 1) {
    cat("\n\n### Overall separate \n")
    for (target_type in target_types) {
      p1 <- plot_function(data, target_type = target_type, horizon = horizon)
      check_print(p1)
    }
  }

  # by location
  for (location in locations) {

    cat("\n\n###", location, "\n\n")
    # cat("\n\n #### Overall \n\n")
    # for (target_type in target_types) {
    #   p1 <- plot_function(data, target_type = target_type, 
                            location = location, horizon = horizon)
    #   check_print(p1)
    # }

    for (target_type in target_types) {
      p1 <- plot_function(data, target_type = target_type,
                          location = location, horizon = horizon)
      check_print(p1)
    }
  }
}

```

# Interval Score Components  {.tabset .tabset-pill}

Weighted interval score broken down into its components "overprediction penalty",
"underprediction penalty" and "sharpness"

``` {r wis-components, fig.width = 12, results = 'asis', eval = TRUE, message = FALSE}
# overall summary
plot_wis_components_helper <-  function(data, target_type = NULL, location = NULL,
                                        forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)

  if (nrow(df) == 0) {
    return(NULL)
  }
  
  scores <- scoringutils::eval_forecasts(df,
                                         summarise_by = "model")

  if (nrow(scores) == 0) {
    return(NULL)
  }
  
  plot <- scoringutils::wis_components(scores) +
    ggplot2::ggtitle(paste("WIS componentents - Incident", target_type,
                           "forecast"))
  return(plot)
}

plot_function <- plot_wis_components_helper


# replace by template ##########################################################
cat("\n\n## Overall \n\n")
pp <- plot_function(data)
check_print(pp)

if (length(target_types) > 1) {
  cat("\n\n## Overall separate \n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}




# by location
for (location in locations) {
  cat("\n\n## ", location, "{.tabset}\n\n")
  
  cat("\n\n### Overall \n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type, location = location)
    check_print(p1)
  }
  
  for (forecast_date in forecast_dates) {
    cat(paste0("\n\n### ", forecast_date), "\n\n")
    for (target_type in target_types) {
      p1 <- plot_function(data, target_type = target_type,
                          location = location,
                          forecast_date = forecast_date)
      check_print(p1)
    }
  }
}
# replace by template ##########################################################

```


# Calibration {.tabset}

## Interval Coverage {.tabset .tabset-pills}

``` {r fig.width = 12, results = 'asis', eval = TRUE}
plot_interval_coverage_helper <- function(data, target_type = NULL, location = NULL,
                                          forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)
  if (nrow(df) == 0) {
    return(NULL)
  }
  scores <- scoringutils::eval_forecasts(df,
                                         summarise_by = c("model", "range"))
  if (nrow(scores) == 0) {
    return(NULL)
  }
  plot <- scoringutils::interval_coverage(scores) +
    ggplot2::ggtitle(paste("Interval coverage - Incident", target_type, "forecasst"))
  return(plot)
}

plot_function <- plot_interval_coverage_helper


# replace by template ##########################################################
cat("\n\n### Overall \n\n")
pp <- plot_function(data)
check_print(pp)

if (length(target_types) > 1) {
  cat("\n\n### Overall separate \n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}

# # by location
# for (location in locations) {
#   cat("\n\n###", location, "{.tabset}\n\n")
#   cat("\n\n #### Overall \n\n")
#   for (target_type in target_types) {
#     p1 <- plot_function(data, target_type = target_type, location = location)
#     check_print(p1)
#   }
#   for (forecast_date in forecast_dates) {
#     cat(paste0("\n\n#### ", forecast_date), "\n\n")
#     for (target_type in target_types) {
#       p1 <- plot_function(data, target_type = target_type,
#                           location = location,
#                           forecast_date = forecast_date)
#       check_print(p1)
#     }
#   }
# }
# replace by template ##########################################################
```

Plot of the percentage of true values captured by each prediction interval

## Quantile Coverage {.tabset .tabset-pills}

``` {r fig.width = 12, results = 'asis', eval = TRUE}
plot_quantile_coverage_helper <- function(data, target_type = NULL, location = NULL,
                                          forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)

  if (nrow(df) == 0) {
    return(NULL)
  }
  scores <- scoringutils::eval_forecasts(df,
                                         summarise_by = c("model", "quantile"))
  if (nrow(scores) == 0) {
    return(NULL)
  }

  plot <- scoringutils::quantile_coverage(scores) +
    ggplot2::ggtitle(paste("Interval coverage - Incident", target_type, "forecast"))
  return(plot)
}

plot_function <- plot_quantile_coverage_helper


# replace by template ##########################################################
cat("\n\n### Overall \n\n")
pp <- plot_function(data)
print(pp)
cat("\n\n")

if (length(target_types) > 1) {
  cat("\n\n### Overall separate \n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}

# by location
# for (location in locations) {
#   cat("\n\n###", location, "{.tabset}\n\n")
#   cat("\n\n #### Overall \n\n")
#   for (target_type in target_types) {
#     p1 <- plot_function(data, target_type = target_type, location = location)
#     check_print(p1)
#   }
#   for (forecast_date in forecast_dates) {
#     cat(paste0("\n\n#### ", forecast_date), "\n\n")
#     for (target_type in target_types) {
#       p1 <- plot_function(data, target_type = target_type,
#                           location = location,
#                           forecast_date = forecast_date)
#       check_print(p1)
#     }
#   }
# }

```

Plot of the percentage of true values below each predictive quantile level


# Bias {.tabset}

``` {r bias, fig.width = 12, results = 'asis', eval = TRUE}
plot_bias_helper <- function(data, target_type = NULL, location = NULL,
                             forecast_date = NULL, horizon = NULL) {

  df <- filter_helper(data, location = location,
                      target_type = target_type,
                      forecast_date = forecast_date,
                      horizon = horizon)

  if (nrow(df) == 0) {
    return(NULL)
  }
  scores <- scoringutils::eval_forecasts(df,
                                         summarise_by = c("model", "location_name"))
  if (nrow(scores) == 0) {
    return(NULL)
  }

  plot <- scoringutils::score_heatmap(scores,
                                      metric = "bias", x = "location_name") +
    ggplot2::labs(x = "Location") +
    ggplot2::ggtitle(paste("Bias - Incident", target_type))
  return(plot)
}

plot_function <- plot_bias_helper


# replace by template ##########################################################
cat("\n\n## Overall \n")
pp <- plot_function(data)
check_print(pp)

if (length(target_types) > 1) {
  cat("\n\n## Overall separate \n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type)
    check_print(p1)
  }
}


cat("\n\n## Separate by date {.tabset} \n\n")
for (forecast_date in forecast_dates) {
  cat(paste0("\n\n### ", forecast_date), "\n\n")
  for (target_type in target_types) {
    p1 <- plot_function(data, target_type = target_type,,
                        forecast_date = forecast_date)
    check_print(p1)
  }
}
```




