---
title: "Handling missing forecasts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling missing forecasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>"
)
data.table::setDTthreads(2)
```

When models forecast different subsets of targets, score summaries become misleading.
A model that only forecasts easy targets will appear to perform better than one that attempts harder targets too.
To get fair comparisons, you need to either restrict to a common set of targets or impute the missing scores.

## Diagnosing missingness

`get_forecast_counts()` shows which models cover which targets.

```{r counts}
library(scoringutils)
fc <- as_forecast_quantile(example_quantile)
get_forecast_counts(fc, by = c("model", "target_type"))
```

Here `UMass-MechBayes` has no case forecasts and `epiforecasts-EpiNow2` has fewer death forecasts than the other models.

## Filtering to a common set of targets

`filter_scores()` removes scores for target combinations not shared across models.
The default strategy, `filter_to_intersection()`, keeps only targets covered by all models.

```{r score}
scores <- score(fc)
```

```{r filter}
scores_filtered <- filter_scores(scores)
```

You can relax the requirement with `min_coverage` to keep targets covered by at least a given proportion of models.

```{r filter-relaxed}
scores_relaxed <- filter_scores(
  scores,
  strategy = filter_to_intersection(
    min_coverage = 0.75
  )
)
```

The `include` argument restricts to targets covered by specific models.

```{r filter-include}
scores_baseline <- filter_scores(
  scores,
  strategy = filter_to_intersection(
    include = "EuroCOVIDhub-baseline"
  )
)
```

## Imputing missing scores

Instead of dropping data, `impute_missing_scores()` fills in scores for missing target combinations.
Four built-in strategies are available.

- `impute_worst_score()` fills with the worst (maximum) observed score for each target.
- `impute_mean_score()` fills with the mean observed score.
- `impute_na_score()` fills with `NA`.
- `impute_model_score(model)` fills with the scores from a named reference model.

```{r impute}
scores_imputed <- impute_missing_scores(
  scores,
  strategy = impute_worst_score()
)
```

Imputed rows are marked with `.imputed = TRUE`, so you can inspect what was added.

```{r imputed-rows}
scores_imputed[scores_imputed$.imputed == TRUE, 1:5]
```

## Recommended workflow

Filter first to remove targets with very low coverage, then impute the remainder.

```{r pipeline}
result <- scores |>
  filter_scores(
    strategy = filter_to_intersection(
      min_coverage = 0.5
    )
  ) |>
  impute_missing_scores(
    strategy = impute_worst_score()
  )
```

You can then summarise as usual with `summarise_scores()`.

```{r summarise}
summarise_scores(result, by = "model")
```
