---
title: "Handling missing forecasts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling missing forecasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>"
)
data.table::setDTthreads(2)
```

When comparing forecast models, not all models will have made predictions for every target.
Naively averaging scores across different sets of targets produces misleading summaries.
This vignette walks through diagnosing which targets each model covers, scoring the forecasts, and then adjusting the scores using filtering and imputation so that summaries are comparable.

## Diagnosing missingness

Before adjusting scores it is worth understanding the pattern of missingness.
Models may have different coverage for legitimate reasons (a model may only forecast deaths, not cases) or because of operational failures (a missed submission deadline).

`get_forecast_counts()` tabulates how many forecasts each model has, grouped by the columns you choose.

```{r counts}
library(scoringutils)
fc <- as_forecast_quantile(example_quantile)
get_forecast_counts(fc, by = c("model", "target_type"))
```

`UMass-MechBayes` does not forecast cases at all, and `epiforecasts-EpiNow2` has fewer death forecasts than the other models.

To see exactly which death targets `epiforecasts-EpiNow2` is missing, we can request counts at a finer level and filter to zero-count rows.

```{r missing-detail}
death_counts <- get_forecast_counts(
  fc,
  by = c("model", "target_type", "location", "target_end_date")
)
death_counts[
  model == "epiforecasts-EpiNow2" &
    target_type == "Deaths" &
    count == 0
]
```

Note that `get_pairwise_comparisons()` has long handled missingness internally, restricting each pair of models to their shared set of targets before comparing.
The functions below give you explicit control over how to handle missingness when computing score summaries.

## Scoring

We score all forecasts first, then adjust the resulting scores table.

```{r score}
scores <- score(fc)
```

A naive summary averages over different numbers of targets per model, making direct comparison unreliable.

```{r naive-summary}
summarise_scores(scores, by = "model")
```

The sections below show two approaches to fixing this: filtering scores to a common set of targets, and imputing scores for missing targets.

## Filtering to a common set of targets

`filter_scores()` removes scores for target combinations not shared across models.
This is appropriate when a model legitimately does not cover certain targets and you want to compare only on shared ground.

The default strategy, `filter_to_intersection()`, keeps only targets covered by **all** models.

```{r filter-default}
scores_filtered <- filter_scores(scores)
summarise_scores(scores_filtered, by = "model")
```

This drops all case targets (since `UMass-MechBayes` has no case forecasts) and the death targets `epiforecasts-EpiNow2` missed.

### Relaxing with min_coverage

You can relax the requirement with `min_coverage`, keeping targets covered by at least a given proportion of models.
With four models, `min_coverage = 0.75` requires coverage by at least three.

```{r filter-relaxed}
scores_relaxed <- filter_scores(
  scores,
  strategy = filter_to_intersection(min_coverage = 0.75)
)
summarise_scores(scores_relaxed, by = "model")
```

This retains the case targets (covered by three of four models) while still dropping any target covered by fewer.

### Filtering to a specific model's targets

The `include` argument restricts to targets covered by named models.
For example, to evaluate all models only on the targets `epiforecasts-EpiNow2` covered:

```{r filter-include}
scores_epinow2 <- filter_scores(
  scores,
  strategy = filter_to_intersection(
    include = "epiforecasts-EpiNow2"
  )
)
summarise_scores(scores_epinow2, by = "model")
```

This keeps both case and death targets where EpiNow2 submitted forecasts, but `UMass-MechBayes` will still be missing case scores in the result.

## Imputing missing scores

Instead of dropping data, `impute_missing_scores()` fills in scores for target combinations a model did not cover.
Imputed rows are marked with `.imputed = TRUE` so they can be identified later.

### Worst score

The simplest penalty: fill each missing score with the worst (maximum) observed score for that target.

```{r impute-worst}
scores_worst <- impute_missing_scores(
  scores,
  strategy = impute_worst_score()
)
summarise_scores(scores_worst, by = "model")
```

You can inspect which rows were imputed:

```{r imputed-check}
scores_worst[(.imputed)]
```

### Mean score

A less severe penalty: fill with the mean score across models that did forecast each target.

```{r impute-mean}
scores_mean <- impute_missing_scores(
  scores,
  strategy = impute_mean_score()
)
summarise_scores(scores_mean, by = "model")
```

### NA

Fill with `NA`, which propagates through summaries unless explicitly handled.

```{r impute-na}
scores_na <- impute_missing_scores(
  scores,
  strategy = impute_na_score()
)
summarise_scores(scores_na, by = "model")
```

### Reference model

Fill with the scores of a named baseline model, treating a missing forecast as performing no better than the baseline.

```{r impute-model}
scores_ref <- impute_missing_scores(
  scores,
  strategy = impute_model_score("EuroCOVIDhub-baseline")
)
summarise_scores(scores_ref, by = "model")
```

## Combining filter and impute

When you want to focus on a specific model's targets but still need complete scores for all models, combine filtering with imputation.
For example, to evaluate on the targets `epiforecasts-EpiNow2` covered and impute scores for models that are missing forecasts within that set:

```{r pipeline}
result <- scores |>
  filter_scores(
    strategy = filter_to_intersection(
      include = "epiforecasts-EpiNow2"
    )
  ) |>
  impute_missing_scores(
    strategy = impute_worst_score()
  )
summarise_scores(result, by = "model")
```
