---
title: "Handling missing forecasts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling missing forecasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>"
)
data.table::setDTthreads(2)
```

When comparing forecast models, not all models will have made predictions for every target.
Naively averaging scores across different sets of targets can produce misleading summaries.
This vignette walks through diagnosing which targets each model covers, scoring the forecasts, and then adjusting the scores using filtering and imputation so that summaries are comparable.

The approaches here were initially inspired by [Kim, Ray & Reich (2026)](https://doi.org/10.1016/j.ijforecast.2025.12.006), who discuss the importance of handling missing forecasts when evaluating model contributions beyond simple leaderboard rankings.

## Diagnosing missingness

Before adjusting scores we should aim to understand the patterns of potential missingness.
Models may have different coverage for legitimate reasons (a model may only forecast deaths, not cases) or because of operational failures (a missed submission deadline).

`get_forecast_counts()` tabulates how many forecasts each model has, grouped by the columns you choose.

```{r counts}
library(scoringutils)
fc <- as_forecast_quantile(example_quantile)
get_forecast_counts(fc, by = c("model", "target_type"))
```

`UMass-MechBayes` does not forecast cases at all, and `epiforecasts-EpiNow2` has fewer death forecasts than the other models.

To see exactly which death targets `epiforecasts-EpiNow2` is missing, we can request counts at a finer level and filter to zero-count rows.

```{r missing-detail}
death_counts <- get_forecast_counts(
  fc,
  by = c("model", "target_type", "location", "target_end_date")
)
death_counts[
  model == "epiforecasts-EpiNow2" &
    target_type == "Deaths" &
    count == 0
]
```

Note that `get_pairwise_comparisons()` handles missingness internally, restricting each pair of models to their shared set of targets before comparing.
The functions below give you explicit control over how to handle missingness when computing score summaries.

## Scoring

We score all forecasts first, then adjust the resulting scores table.
Both `filter_scores()` and `impute_missing_scores()` accept a `compare` argument (default `"model"`) that specifies which column identifies the units being compared.

```{r score}
scores <- score(fc)
```

A naive summary averages over different numbers of targets per model, which can make direct comparison misleading.

```{r naive-summary}
summarise_scores(scores, by = "model")
```

The sections below show two approaches to addressing this: filtering scores to a common set of targets, and imputing scores for missing targets.

## Filtering to a common set of targets

`filter_scores()` removes scores based on the supplied strategy.
This can be appropriate when a model legitimately does not cover certain targets and you want to compare only on shared ground.

The default strategy, `filter_to_intersection()`, keeps only targets covered by **all** models.

```{r filter-default}
scores_filtered <- filter_scores(scores)
summarise_scores(scores_filtered, by = "model")
```

This drops all case targets (since `UMass-MechBayes` has no case forecasts) and the death targets `epiforecasts-EpiNow2` missed.

### Requiring partial coverage

The default requires every model to cover a target for it to be kept.
The `min_coverage` argument relaxes this by keeping targets covered by at least a given proportion of models.
With four models, `min_coverage = 0.75` requires coverage by at least three.

```{r filter-relaxed}
scores_relaxed <- filter_scores(
  scores,
  strategy = filter_to_intersection(min_coverage = 0.75)
)
summarise_scores(scores_relaxed, by = "model")
```

In this example, case targets are covered by three of four models and death targets by three or four, so `min_coverage = 0.75` retains all targets.
The default (`min_coverage = 1`) is stricter and drops all case targets because `UMass-MechBayes` has no case forecasts.
Between 0.75 and 1.0 no intermediate threshold changes the result here because no target is covered by exactly three out of four models while being missing for one that isn't `UMass-MechBayes`.

### Filtering to a specific model's targets

The `include` argument restricts to targets covered by named models.
For example, to evaluate all models only on the targets `epiforecasts-EpiNow2` covered:

```{r filter-include}
scores_epinow2 <- filter_scores(
  scores,
  strategy = filter_to_intersection(
    include = "epiforecasts-EpiNow2"
  )
)
summarise_scores(scores_epinow2, by = "model")
```

This keeps both case and death targets where EpiNow2 submitted forecasts, but `UMass-MechBayes` will still be missing case scores in the result.

## Imputing missing scores

Instead of dropping data, `impute_missing_scores()` fills in scores for target combinations a model did not cover.
Imputed rows are marked with `.imputed = TRUE` so they can be identified later.

### NA

The simplest option: fill missing scores with `NA`.
This preserves the structure of the data without making assumptions about what the score would have been.
`NA` values propagate through summaries, so this can also serve as a diagnostic check to confirm where missingness exists.

```{r impute-na}
scores_na <- impute_missing_scores(
  scores,
  strategy = impute_na_score()
)
summarise_scores(scores_na, by = "model")
```

### Worst score

Fill each missing score with the worst (maximum) observed score for that target across all models.
This penalises models most heavily for missing targets.

```{r impute-worst}
scores_worst <- impute_missing_scores(
  scores,
  strategy = impute_worst_score()
)
summarise_scores(scores_worst, by = "model")
```

We can check that the imputed rows match the models and targets we identified as missing earlier.

```{r imputed-check}
scores_worst[
  (.imputed),
  .(n_imputed = .N),
  by = c("model", "target_type")
]
```

### Reference model

Fill with the scores of a named baseline model, treating a missing forecast as performing no better than that baseline.

```{r impute-model}
scores_ref <- impute_missing_scores(
  scores,
  strategy = impute_model_score("EuroCOVIDhub-baseline")
)
summarise_scores(scores_ref, by = "model")
```

This is a reasonable default when a suitable baseline exists, though more research is needed on best practice for choosing the reference model and understanding the impact of this choice.

### Mean score

Fill with the mean score across models that did forecast each target.
This is the least severe penalty as it assigns the average performance, which may be close to the ensemble performance.

```{r impute-mean}
scores_mean <- impute_missing_scores(
  scores,
  strategy = impute_mean_score()
)
summarise_scores(scores_mean, by = "model")
```

## Combining filter and impute

Consider combining filtering and imputation when you want to focus on a specific model's targets but still need complete scores for all models.
For example, to evaluate on the targets `epiforecasts-EpiNow2` covered and then impute scores for models that are missing forecasts within that set:

```{r pipeline}
result <- scores |>
  filter_scores(
    strategy = filter_to_intersection(
      include = "epiforecasts-EpiNow2"
    )
  ) |>
  impute_missing_scores(
    strategy = impute_worst_score()
  )
summarise_scores(result, by = "model")
```
