---
title: "scoringutils"
author: "Nikos Bosse"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scoringutils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction and Overview of Functionality

The `scoringutils` package provides a collection of metrics and proper scoring rules
that make it simple to score forecasts against the true observed values.
Predictions can either be automatically scored from a `data.frame` using the function `eval_forecasts`. Alternatively, evaluation metrics can be accessed directly using lower level functions within a vector/matrix framework. 

Predictions can be handled in various formats: `scoringutils` can handle probabilistic forecasts in either a sample based or a quantile based format. For more detail on the expected input formats please see below. True values can be integer, continuous or binary. 

In addition to automatic scoring, `scoringutils` offers a variety of plots and visualisations. 

# Scoring Forecasts Automatically

Most of the time, the `eval_forecasts` function will be able to do the entire evaluation for you. The idea is simple, yet flexible. 

All you need to do is to pass in a `data.frame` that has a column called `prediction` and one called `true_value`. Depending on the exact input format, additional columns like `sample`, `quantile` or `range` and `boundary` are needed. Additional columns may be present to indicate a grouping of forecasts. For example, we could have forecasts made by different models in various locations at different time points, each for several weeks into the future. In this case, we would have additional columns called for example `model`, `date`, `forecast_date`, `forecast_horizon` and `location`. 

Using the `by` argument you need to specify the *unit of a single forecast*. In this example here we would set `by = c("model", "date", "forecast_date", "forecast_horizon", "location")` (note: if we want to be pedantic, there is a small duplication as the information of "date" is already included in the combination of "forecast_date" and "forecast_horizon". But as long as there isn't some weird shift, this doesn't matter for the purpose of grouping our observations). If you don't specify `by` (i.e. `by = NULL`), `scoringutils` will automatically use all appropriate present columns. Note that you don't need to include columns such as `quantile` or `sample` in the `by` argument, as several quantiles / samples make up one forecast. 

Using the `summarise_by` argument you can now choose categories to aggregate over. If you were only interested in scores for the different models, you would specify `summarise_by = c("model")`. If you wanted to have scores for every model in every location, you would need to specify `summarise_by = c("model", "location")`. If you wanted to have one score per quantile or one per prediction interval range, you could specify something like `summarise_by = c("model", "quantile")` or `summarise_by = c("model", "quantile", "range")` (note again that some information is duplicated in quantile and range, but this doesn't really matter for grouping purposes). When aggregating, `eval_forecasts` takes the mean according to the group defined in `summarise_by` (i.e. in this example, if `summarise_by = c("model", "location")`, scores will be averaged over all forecast dates, forecast horizons and quantiles to yield one score per model and location). In addition to the mean, you can also obtain the standard deviation of the scores over which you average or any desired quantile (e.g. the median in addition to the mean) by specifying `sd = TRUE` and `quantiles = c(0.5)`. 


## Example Evaluation 

Here is an example of an evaluation using toy data:

```{r}
library(scoringutils)
library(data.table)
```

```{r}
data <- scoringutils::quantile_example_data_plain
print(data, 3, 3)
scoringutils::eval_forecasts(data, 
                             summarise_by = c("model", "quantile", "range"))
```

Using an appropriate level of summary, we can easily use the output for visualisation. The `scoringutils` package offers some built-in functions to help get a sense of the data

```{r}
scoringutils::plot_predictions(data, x = "id", range = c(0, 90), 
                               facet_formula = ~ model)
```

(The data is just randomly generated values. We plan to add real example data to make these illustrations more useful in the future)

```{r}
scores <- scoringutils::eval_forecasts(data, 
                             summarise_by = c("model"))
scoringutils::score_table(scores)
```

Given this level of aggregation, not all metrics may make sense. In this case, for example, averaging over different quantiles to compute quantile coverage does not make much sense. If you like, you can select specific metrics for the visualisation. 

Let us look at calibration: 

```{r out.width="50%", fig.show="hold"}
scores <- scoringutils::eval_forecasts(data, 
                             summarise_by = c("model", "range", "quantile"))
scoringutils::interval_coverage(scores) + 
  ggplot2::ggtitle("Interval Coverage")

scoringutils::quantile_coverage(scores) + 
  ggplot2::ggtitle("Quantile Coverage")
```

Let us look at the individual components of the weighted interval score: 

```{r}
scores <- scoringutils::eval_forecasts(data, 
                             summarise_by = c("model"))
scoringutils::wis_components(scores)
```

We can also look at contributions to different metrics by range: 

```{r}
scores <- scoringutils::eval_forecasts(data, 
                             summarise_by = c("model", "range"))
scoringutils::range_plot(scores, y = "interval_score")
```

We can also visualise metrics using a heatmap: 

```{r}
scores <- scoringutils::eval_forecasts(data, 
                             summarise_by = c("model", "horizon"))
scores[, horizon := as.factor(horizon)]
scoringutils::score_heatmap(scores, 
                            x = "horizon", metric = "bias")
```


### Expected Input Formats

The `eval_forecasts` function is designed to work with various different input formats. The following formats are currently supported: 

quantile forecasts in either a plain quantile format or in a format that specifies interval ranges and the boundary of a given interval range. 

``` {r}
print(scoringutils::quantile_example_data_plain, 3, 3)
print(scoringutils::quantile_example_data_long, 3, 3)
```

sample based format with either continuous or integer values

``` {r}
print(scoringutils::integer_example_data, 3, 3)
print(scoringutils::continuous_example_data, 3, 3)
```

forecasts in a binary format:

``` {r}
print(scoringutils::binary_example_data, 3, 3)
```

It also offers functionality to convert between these formats. For more information have a look at the documentation of the following functions: 
``` {r eval=FALSE}
scoringutils::sample_to_quantile() # convert from sample based to quantile format
scoringutils::range_to_quantile() # convert from range format to plain quantile
scoringutils::quantile_to_range() # convert the other way round
scoringutils::quantile_to_long() # convert range based format from wide to long
scoringutils::quantile_to_wide() # convert the other way round
```


# Scoring Forecasts Directly

A variety of metrics and scoring rules can also be accessed directly through 
the `scoringutils` package. 

The following gives an overview of (most of) the implemented metrics. 

## Bias

The function `bias` determines bias from predictive Monte-Carlo samples, 
automatically recognising whether forecasts are continuous or
integer valued. 

For continuous forecasts, Bias is measured as
$$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t))$$

where $P_t$ is the empirical cumulative distribution function of the
prediction for the true value $x_t$. Computationally, $P_t (x_t)$ is
just calculated as the fraction of predictive samples for $x_t$
that are smaller than $x_t$.

For integer valued forecasts, Bias is measured as

$$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1))$$

to adjust for the integer nature of the forecasts. In both cases, Bias can 
assume values between -1 and 1 and is 0 ideally.

```{r}
## integer valued forecasts
true_values <- rpois(30, lambda = 1:30)
predictions <- replicate(200, rpois(n = 30, lambda = 1:30))
bias(true_values, predictions)

## continuous forecasts
true_values <- rnorm(30, mean = 1:30)
predictions <- replicate(200, rnorm(30, mean = 1:30))
bias(true_values, predictions)
```


## Sharpness
Sharpness is the ability of the model to generate predictions within a
narrow range. It is a data-independent measure, and is purely a feature
of the forecasts themselves.

Shaprness of predictive samples corresponding to one single true value is
measured as the normalised median of the absolute deviation from
the median of the predictive samples. For details, see `?stats::mad`

```{r}
predictions <- replicate(200, rpois(n = 30, lambda = 1:30))
sharpness(predictions)
```

## Calibration

Calibration or reliability of forecasts is the ability of a model to
correctly identify its own uncertainty in making predictions. In a model
with perfect calibration, the observed data at each time point look as if
they came from the predictive probability distribution at that time.

Equivalently, one can inspect the probability integral transform of the
predictive distribution at time t,

$$u_t = F_t (x_t)$$

where $x_t$ is the observed data point at time $t \text{ in } t_1, …, t_n$,
n being the number of forecasts, and $F_t$ is the (continuous) predictive
cumulative probability distribution at time t. If the true probability
distribution of outcomes at time t is $G_t$ then the forecasts $F_t$ are
said to be ideal if $F_t = G_t$ at all times $t$. In that case, the
probabilities ut are distributed uniformly.

In the case of discrete outcomes such as incidence counts,
the PIT is no longer uniform even when forecasts are ideal.
In that case a randomised PIT can be used instead:

$$u_t = P_t(k_t) + v \cdot (P_t(k_t) - P_t(k_t - 1) )$$

where $k_t$ is the observed count, $P_t(x)$ is the predictive
cumulative probability of observing incidence $k$ at time $t$,
$P_t (-1) = 0$ by definition and $v$ is standard uniform and independent
of $k$. If $P_t$ is the true cumulative
probability distribution, then $u_t$ is standard uniform.

The function checks whether integer or continuous forecasts were provided.
It then applies the (randomised) probability integral and tests
the values $u_t$ for uniformity using the
Anderson-Darling test.

As a rule of thumb, there is no evidence to suggest a forecasting model is
miscalibrated if the p-value found was greater than a threshold of $p >= 0.1$,
some evidence that it was miscalibrated if $0.01 < p < 0.1$, and good
evidence that it was miscalibrated if $p <= 0.01$.
In this context it should be noted, though, that uniformity of the
PIT is a necessary but not sufficient condition of calibration. It should
als be noted that the test only works given sufficient samples, otherwise the 
Null hypothesis will often be rejected outright. 


## Continuous Ranked Probability Score (CRPS)
Wrapper around the `crps_sample` function from the
`scoringRules` package. For more information look at the manuals from the
`scoringRules` package. The function can be used for continuous as well as 
integer valued forecasts. Smaller values are better. 

```{r}
true_values <- rpois(30, lambda = 1:30)
predictions <- replicate(200, rpois(n = 30, lambda = 1:30))
crps(true_values, predictions)
```



## Dawid-Sebastiani Score (DSS)
Wrapper around the `dss_sample` function from the
`scoringRules` package. For more information look at the manuals from the
`scoringRules` package. The function can be used for continuous as well as 
integer valued forecasts. Smaller values are better. 

```{r}
true_values <- rpois(30, lambda = 1:30)
predictions <- replicate(200, rpois(n = 30, lambda = 1:30))
dss(true_values, predictions)
```

## Log Score
Wrapper around the `log_sample` function from the
`scoringRules` package. For more information look at the manuals from the
`scoringRules` package. The function should not be used for integer valued 
forecasts. While Log Scores are in principle possible for integer valued 
foreasts they require a kernel density estimate which is not well defined 
for discrete values. Smaller values are better. 

```{r}
true_values <- rnorm(30, mean = 1:30)
predictions <- replicate(200, rnorm(n = 30, mean = 1:30))
logs(true_values, predictions)
```

## Brier Score
The Brier score is a proper score rule that assesses the accuracy of
probabilistic binary predictions. The outcomes can be either 0 or 1,
the predictions must be a probability that the true outcome will be 1.

The Brier Score is then computed as the mean squared error between the
probabilistic prediction and the true outcome.

$$\text{Brier_Score} = \frac{1}{N} \sum_{t = 1}^{n} (\text{prediction}_t - \text{outcome}_t)^2$$


```{r}
true_values <- sample(c(0,1), size = 30, replace = TRUE)
predictions <- runif(n = 30, min = 0, max = 1)

brier_score(true_values, predictions)
```

## Interval Score
The Interval Score is a Proper Scoring Rule to score quantile predictions,
following Gneiting and Raftery (2007). Smaller values are better.

The score is computed as

$$ \text{score} = (\text{upper} - \text{lower}) + \\
\frac{2}{\alpha} \cdot (\text{lower} - \text{true_value}) \cdot 1(\text{true_values} < \text{lower}) + \\
\frac{2}{\alpha} \cdot (\text{true_value} - \text{upper}) \cdot
1(\text{true_value} > \text{upper})$$


where $1()$ is the indicator function and $\alpha$ is the decimal value that
indicates how much is outside the prediction interval.
To improve usability, the user is asked to provide an interval range in
percentage terms, i.e. interval_range = 90 (percent) for a 90 percent
prediction interval. Correspondingly, the user would have to provide the
5\% and 95\% quantiles (the corresponding alpha would then be 0.1).
No specific distribution is assumed,
but the range has to be symmetric (i.e you can't use the 0.1 quantile
as the lower bound and the 0.7 quantile as the upper). 
Setting `weigh = TRUE` will weigh the score by $\frac{\alpha}{2}$ such that 
the Interval Score converges to the CRPS for increasing number of quantiles. 


```{r}
true_values <- rnorm(30, mean = 1:30)
interval_range <- 90
alpha <- (100 - interval_range) / 100
lower <- qnorm(alpha/2, rnorm(30, mean = 1:30))
upper <- qnorm((1- alpha/2), rnorm(30, mean = 1:30))

interval_score(true_values = true_values,
               lower = lower,
               upper = upper,
               interval_range = interval_range)
```

