---
title: "Visualisations"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Visualisations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This Vignette provides a preliminary overview of visualisations for forecasts and scores. Section [Visualising forecasts] gives an overview of how to visualise forecasts. Section [Visualising scores] provides an overview of how to visualise scores. Section [Old visualisations] provides the code for visualisations that were previously part of `scoringutils` but have now been removed.

The example data used is the data shipped with the `scoringutils` package. It consists of short-term forecasts of COVID-19 cases and deaths for several European countries, submitted to the European COVID-19 Forecast Hub. Every week, forecasters submitted 1 to 4 week ahead forecasts (the "forecast horizon", here restricted to 1 to 3 week ahead). The dates for which a forecast was made are called `target_end_date`. See e.g. `?example_quantile` for more information.

```{r setup, message=FALSE}
library(scoringutils)
library(ggplot2)
library(tidyr)
library(dplyr)
```

# Visualising forecasts

## Timeseries plot with only one horizon (visualisation of all 2 week ahead forecasts and observed values)

This plot shows a visualisation of all available two week ahead forecasts and observed data for Germany. 

```{r}
two_week_forecasts <- example_quantile %>%
  filter(model %in% c(NA,"EuroCOVIDhub-ensemble"), 
         target_type == "Cases", 
         location == "DE", 
         horizon == 2)

two_week_forecasts %>%
  pivot_wider(names_from = quantile_level, values_from = c("predicted")) %>%
  ggplot(aes(x = target_end_date)) + 
  geom_ribbon(aes(ymin = `0.025`, ymax = `0.975`), fill = "skyblue2", alpha = 0.3) + 
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), fill = "skyblue2", alpha = 0.3) + 
  geom_line(aes(y = `0.5`), linewidth = 1, colour = "skyblue2") + 
  geom_line(aes(y = observed), linewidth = 0.5) +
  geom_point(aes(y = observed)) +
  theme_scoringutils()
```


## Timeseries plot with only one forecast date, but several forecast horizons

```{r}
forecasts_one_date <- example_quantile %>%
  filter(model %in% c(NA,"EuroCOVIDhub-ensemble"), 
         target_type == "Cases", 
         forecast_date == "2021-06-14",
         location == "DE")

forecasts_one_date %>%
  pivot_wider(names_from = quantile_level, values_from = c("predicted")) %>%
  ggplot(aes(x = target_end_date)) + 
  geom_ribbon(aes(ymin = `0.025`, ymax = `0.975`), fill = "skyblue2", alpha = 0.3) +
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), fill = "skyblue2", alpha = 0.3) +
  geom_line(aes(y = `0.5`), linewidth = 1, colour = "skyblue2") +
  geom_line(aes(y = observed), linewidth = 0.5) +
  geom_point(aes(y = observed)) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2021-06-14"))), linewidth = 4, colour = "black") + 
  theme_scoringutils()
```



# Visualising scores

To be expanded later. 

# Old visualisations

## Functions `plot_predictions()` and `make_na()`

In previous versions of `scoringutils`, forecasts and observed values could be visualised using the function `plot_predictions()` and its `make_na()` helper function. The following shows the function code first and then an example. 

```{r}
#' @title Plot Predictions vs True Values
#'
#' @description
#' Make a plot of observed and predicted values
#'
#' @param data a data.frame that follows the same specifications outlined in
#' [score()]. To customise your plotting, you can filter your data using the
#' function [make_NA()].
#' @param by character vector with column names that denote categories by which
#' the plot should be stratified. If for example you want to have a facetted
#' plot, this should be a character vector with the columns used in facetting
#' (note that the facetting still needs to be done outside of the function call)
#' @param x character vector of length one that denotes the name of the variable
#' @param interval_range numeric vector indicating the interval ranges to plot.
#' If 0 is included in `interval_range`, the median prediction will be shown.
#' @return ggplot object with a plot of true vs predicted values
#' @importFrom ggplot2 ggplot scale_colour_manual scale_fill_manual theme_light
#' @importFrom ggplot2 facet_wrap facet_grid aes geom_line .data geom_point
#' @importFrom data.table dcast
#' @importFrom ggdist geom_lineribbon
#' @export
#' @examples
#' library(ggplot2)
#' library(magrittr)
#'
#' example_continuous %>%
#'   make_NA (
#'     what = "truth",
#'     target_end_date >= "2021-07-22",
#'     target_end_date < "2021-05-01"
#'   ) %>%
#'   make_NA (
#'     what = "forecast",
#'     model != "EuroCOVIDhub-ensemble",
#'     forecast_date != "2021-06-07"
#'   ) %>%
#'   plot_predictions (
#'     x = "target_end_date",
#'     by = c("target_type", "location"),
#'     interval_range = c(0, 50, 90, 95)
#'   ) +
#'   facet_wrap(~ location + target_type, scales = "free_y") +
#'   aes(fill = model, color = model)
#'
#' example_continuous %>%
#'   make_NA (
#'     what = "truth",
#'     target_end_date >= "2021-07-22",
#'     target_end_date < "2021-05-01"
#'   ) %>%
#'   make_NA (
#'     what = "forecast",
#'     forecast_date != "2021-06-07"
#'   ) %>%
#'   plot_predictions (
#'     x = "target_end_date",
#'     by = c("target_type", "location"),
#'     interval_range = 0
#'   ) +
#'   facet_wrap(~ location + target_type, scales = "free_y") +
#'   aes(fill = model, color = model)

plot_predictions <- function(data,
                             by = NULL,
                             x = "date",
                             interval_range = c(0, 50, 90)) {

  # split truth data and forecasts in order to apply different filtering
  truth_data <- data.table::as.data.table(data)[!is.na(observed)]
  forecasts <- data.table::as.data.table(data)[!is.na(predicted)]

  del_cols <-
    colnames(truth_data)[!(colnames(truth_data) %in% c(by, "observed", x))]

  truth_data <- unique(suppressWarnings(truth_data[, eval(del_cols) := NULL]))

  # find out what type of predictions we have. convert sample based to
  # interval range data

  if (get_forecast_type(data) == "quantile") {
    forecasts <- quantile_to_interval(
      forecasts,
      keep_quantile_col = FALSE
    )
  } else if (get_forecast_type(data) == "sample") {
    # using a scoringutils internal function
    forecasts <- scoringutils:::sample_to_interval_long(
      forecasts,
      interval_range = interval_range,
      keep_quantile_col = FALSE
    )
  }

  # select appropriate boundaries and pivot wider
  select <- forecasts$interval_range %in% setdiff(interval_range, 0)
  intervals <- forecasts[select, ]

  # delete quantile column in intervals if present. This is important for
  # pivoting
  if ("quantile_level" %in% names(intervals)) {
    intervals[, quantile_level := NULL]
  }

  plot <- ggplot(data = data, aes(x = .data[[x]])) +
    theme_scoringutils() +
    ylab("True and predicted values")

  if (nrow(intervals) != 0) {
    # pivot wider and convert range to a factor
    intervals <- data.table::dcast(intervals, ... ~ boundary,
                                   value.var = "predicted")

    # only plot interval ranges if there are interval ranges to plot
    plot <- plot +
      ggdist::geom_lineribbon(
        data = intervals,
        aes(
          ymin = lower, ymax = upper,
          # We use the fill_ramp aesthetic for this instead of the default fill
          # because we want to keep fill to be able to use it for other
          # variables
          fill_ramp = factor(
            interval_range,
            levels = sort(unique(interval_range), decreasing = TRUE)
          )
        ),
        lwd = 0.4
      ) +
      ggdist::scale_fill_ramp_discrete(
        name = "interval_range",
        # range argument was added to make sure that the line for the median
        # and the ribbon don't have the same opacity, making the line
        # invisible
        range = c(0.15, 0.75)
      )
  }

  # We could treat this step as part of ggdist::geom_lineribbon() but we treat
  # it separately here to deal with the case when only the median is provided
  # (in which case ggdist::geom_lineribbon() will fail)
  if (0 %in% interval_range) {
    select_median <-
      forecasts$interval_range == 0 & forecasts$boundary == "lower"
    median <- forecasts[select_median]

    if (nrow(median) > 0) {
      plot <- plot +
        geom_line(
          data = median,
          mapping = aes(y = predicted),
          lwd = 0.4
        )
    }
  }

  # add observed values
  if (nrow(truth_data) > 0) {
    plot <- plot +
      geom_point(
        data = truth_data,
        show.legend = FALSE,
        inherit.aes = FALSE,
        aes(x = .data[[x]], y = observed),
        color = "black",
        size = 0.5
      ) +
      geom_line(
        data = truth_data,
        inherit.aes = FALSE,
        show.legend = FALSE,
        aes(x = .data[[x]], y = observed),
        linetype = 1,
        color = "grey40",
        lwd = 0.2
      )
  }

  return(plot)
}
```

`plot_predictions()` does the actual work of producing a plot. The `by` argument is needed so that the user can facet the plot correctly and the user needs to specify all columns relevant for facetting. `make_NA()` represents a form of filtering, but instead of filtering entire rows, the relevant entries in the columns "predicted" or "observed" are made `NA`. This allows the user to filter observations and forecasts independently. 

```{r}
#' @title Make Rows NA in Data for Plotting
#'
#' @description
#' Filters the data and turns values into `NA` before the data gets passed to
#' [plot_predictions()]. The reason to do this is to this is that it allows to
#' 'filter' prediction and truth data separately. Any value that is NA will then
#' be removed in the subsequent call to [plot_predictions()].
#'
#' @inheritParams score
#' @param what character vector that determines which values should be turned
#' into `NA`. If `what = "truth"`, values in the column 'observed' will be
#' turned into `NA`. If `what = "forecast"`, values in the column 'prediction'
#' will be turned into `NA`. If `what = "both"`, values in both column will be
#' turned into `NA`.
#' @param ... logical statements used to filter the data
#' @return A data.table
#' @importFrom rlang enexprs
#' @keywords plotting
#' @export
#'
#' @examples
#' make_NA (
#'     example_continuous,
#'     what = "truth",
#'     target_end_date >= "2021-07-22",
#'     target_end_date < "2021-05-01"
#'   )

make_NA <- function(data = NULL,
                    what = c("truth", "forecast", "both"),
                    ...) {

  stopifnot(is.data.frame(data))
  data <- as.data.table(data)
  what <- match.arg(what)

  # turn ... arguments into expressions
  args <- enexprs(...)

  vars <- NULL
  if (what %in% c("forecast", "both")) {
    vars <- c(vars, "predicted")
  }
  if (what %in% c("truth", "both")) {
    vars <- c(vars, "observed")
  }
  for (expr in args) {
    data <- data[eval(expr), eval(vars) := NA_real_]
  }
  return(data[])
}
```

In the following are a few examples of using the two functions to create a plot using the scoringutils example data. 

```{r}
library(ggplot2)
library(ggdist)
library(data.table)
library(magrittr) #pipe operator
```


Visualising the median forecasts for the example data. The truth data is restricted to a period between 2021-05-01 and 2021-07-22. The forecast data is a forecast from the model "EuroCOVIDhub-ensemble" made on the "2021-06-07". All other data is set to `NA`, effectively removing it from the plot. 

```{r}
median_forecasts <- example_quantile[quantile_level == 0.5]
median_forecasts %>%
  make_NA(what = "truth",
          target_end_date <= '2021-05-01',
          target_end_date > '2021-07-22') %>%
  make_NA(what = "forecast",
          model != "EuroCOVIDhub-ensemble",
          forecast_date != '2021-06-07') %>%
  plot_predictions(
    by = c("location", "target_type"),
    x = "target_end_date",
  ) +
  facet_wrap(location ~ target_type, scales = "free_y")
```

This is the same plot, but with a variety of prediction intervals shown, instead of just the median. 

```{r}
example_quantile %>%
  make_NA(what = "truth",
          target_end_date <= '2021-05-01',
          target_end_date > '2021-07-22') %>%
  make_NA(what = "forecast",
          model != "EuroCOVIDhub-ensemble",
          forecast_date != '2021-06-07') %>%
  plot_predictions(
    by = c("location", "target_type"),
    x = "target_end_date",
    interval_range = c(0, 10, 20, 30, 40, 50, 60)
  ) +
  facet_wrap(location ~ target_type, scales = "free_y")
```

And a similar plot, this time based on continuous forecasts. The predictions are automatically converted to a quantile-based forecasts for plotting.

```{r}
example_continuous %>%
    make_NA(what = "truth",
            target_end_date <= '2021-05-01',
            target_end_date > '2021-07-22') %>%
    make_NA(what = "forecast",
            model != "EuroCOVIDhub-ensemble",
            forecast_date != '2021-06-07') %>%
    plot_predictions(
      by = c("location", "target_type"),
      x = "target_end_date",
      interval_range = c(0, 50, 90, 95)
    ) +
    facet_wrap(location ~ target_type, scales = "free_y")
```

Displaying two forecasts at a time with additional colours: 

```{r}
example_quantile %>%
  make_NA(what = "truth", 
          target_end_date > "2021-07-15",
          target_end_date <= "2021-05-22") %>%
  make_NA(what = "forecast", 
          !(model %in% c("EuroCOVIDhub-ensemble", "EuroCOVIDhub-baseline")),
          forecast_date != "2021-06-28") %>%
  plot_predictions(x = "target_end_date", by = c("target_type", "location")) +
  aes(colour = model, fill = model) +
  facet_wrap(target_type ~ location, ncol = 4, scales = "free_y") +
  labs(x = "Target end date")
```



## Function `plot_interval_ranges()` (formerly `plot_ranges()`)

```{r}
#' @title Plot Metrics by Range of the Prediction Interval
#'
#' @description
#' Visualise the metrics by range, e.g. if you are interested how different
#' interval ranges contribute to the overall interval score, or how
#' sharpness / dispersion changes by range.
#'
#' @param scores A data.frame of scores based on quantile forecasts as
#' produced by [score()] or [summarise_scores()]. Note that "range" must be included
#' in the `by` argument when running [summarise_scores()]
#' @param y The variable from the scores you want to show on the y-Axis.
#' This could be something like "wis" (the default) or "dispersion"
#' @param x The variable from the scores you want to show on the x-Axis.
#' Usually this will be "model"
#' @param colour Character vector of length one used to determine a variable
#' for colouring dots. The Default is "range".
#' @return A ggplot2 object showing a contributions from the three components of
#' the weighted interval score
#' @importFrom ggplot2 ggplot aes aes geom_point geom_line
#' expand_limits theme theme_light element_text scale_color_continuous labs
#' @export
#' @examples
#' library(ggplot2)
#' ex <- data.table::copy(example_quantile)
#' ex$range <- scoringutils:::get_range_from_quantile(ex$quantile)
#' scores <- suppressWarnings(score(as_forecast(ex), metrics = list("wis" = wis)))
#' summarised <- summarise_scores(
#'   scores,
#'   by = c("model", "target_type", "range")
#' )
#' plot_interval_ranges(summarised, x = "model") +
#'   facet_wrap(~target_type, scales = "free")

plot_interval_ranges <- function(scores,
                                 y = "wis",
                                 x = "model",
                                 colour = "range") {
  plot <- ggplot(
    scores,
    aes(
      x = .data[[x]],
      y = .data[[y]],
      colour = .data[[colour]]
    )
  ) +
    geom_point(size = 2) +
    geom_line(aes(group = range),
      colour = "black",
      linewidth = 0.01
    ) +
    scale_color_continuous(low = "steelblue", high = "salmon") +
    theme_scoringutils() +
    expand_limits(y = 0) +
    theme(
      legend.position = "right",
      axis.text.x = element_text(
        angle = 90, vjust = 1,
        hjust = 1
      )
    )

  return(plot)
}
```

The functioniality currently relies on a hack. In previous versions of `scoringutils`, scores were computed per interval range/per quantile. Now, `scoringutils` returns one score per forecast, not per interval range/quantile. We therefore need to add a `range` column, using an internal function `get_range_from_quantile()`. This column will be interpreted as one that defines the unit of a single forecast by `scoringutils`. It also means that we will get a warning about different number of quantile levels for different forecasts (because the 0% prediction interval only has one median forecast, while all other prediction intervals have two (a lower and an upper bound)).

```{r}
range_example <- copy(example_quantile) %>%
  na.omit() %>%
  .[, range := scoringutils:::get_range_from_quantile(quantile_level)]

sum_scores <- range_example %>%
  as_forecast() %>%
  score(metrics = list("wis" = wis, "dispersion" = dispersion)) %>%
  summarise_scores(by = c("model", "target_type", "range")) %>%
  suppressWarnings()

plot_interval_ranges(sum_scores, x = "model") +
  facet_wrap(~target_type, scales = "free")
```

Plotting dispersion instead of WIS: 

```{r}
plot_interval_ranges(sum_scores, y = "dispersion", x = "model") +
    facet_wrap(~target_type, scales = "free_y")
```

